---
title: "Assignment3_Analysis"
output: html_document
---

# Team members
1. [Maria Herdt] [13-704-911]
2. [Justin Sosnoski] [16-726-028]

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

####
# BEGIN: Header
####

# The "pacman" package allows p_load() and p_load_gh() which
# automatically install missing packages and load them
if (!require("pacman")) install.packages("pacman", repos='https://stat.ethz.ch/CRAN/'); library(pacman)

p_load(
   car,        # grab bag of useful functions for NHST
   multcomp,   # for glht (planned comparison, post-hoc test)
   tidyverse,  # collection of the tidyverse packages (this automatically load the following):
   #dplyr,     #   - for data wrangling
   #tibble,    #   - a stricter alternative to data.frame
   #readr,     #   - a stricter alternative to read.csv
   #ggplot2,   #   - for plotting
               # other packages in tidyverse that are non-core
   stringr,    #   - for string functions
   tidyr,      #   - for data tidying
   forcats,    #   - utility functions for working with factor levels
               # extensions of tidyverse
   broom,      #   - for cleaing output from models, e.g., lm()
   cowplot,    #   - adds plot_grid() to put multiple ggplot()'s togeter
   GGally,     #   - adds ggpairs() which is a smarter scatterplot matrix
               # data structure
   GetoptLong, #   - string interpolation. See qq() explanation below
   lubridate,  #   - utility for parsing and performing arithematic on dates 
               # 
               # visualization & interactivity
   Hmisc,      #   - for plotting mean and CI in ggplot
   rafalib,    #   - for imagemat function (visualize contrast codings)
   DT,         #   - for showing data table with navigation/search controls
               # testing:
   assertthat  #   - unit-testing functions
   
)

p_load_gh(
   "eclarke/ggbeeswarm" # beeswarm plot extension for ggplot2
)

# GetoptLong config (string interpolation)
qq.options("code.pattern" = "#\\{CODE\\}") 

# ggplot2 config (plot theme)
myTheme <- theme(
   panel.background = element_blank(), 
   panel.grid.major = element_line(color="lightgrey", size = 0.2)
)

# DT config
options(DT.options = list(pageLength = 10))
options(DT.autoHideNavigation = TRUE)

# Optional: Decimal output readability
# options(digits=2)
##   NOTE: This option might cause output to be printed with rounding. (default value = 7)

####
# END: Header
####
```

# Experimental design and data description

The following dataset is ...

* Four input conditions: 

C01. Computer trackpad (index finger of dominant to touch; baseline)
C02. Mobile phone as touchpad one-handed (phone in dominant hand, thumb to touch)
C03. Mobile phone as touchpad two-handed (phone in non-dominant hand, dominant index finger to touch)
C04. Mobile phone in sway mode

* Three levels of target amplitudes:  (300, 450, 600)

* Three levels of target width:  (30, 45, 60)

* Number of participants: 6

* Number of repetitions per participants: 2 trial blocks

* Total number of results per participant: 4 x 3 x 3 x 2 = 72 results

**Accuracy measures /according to MacKenzie et al (2001)/

TRE	<dbl>	Target re-entry
TAC	<dbl>	Task axis crossing
MDC	<dbl>	Movement direction change
ODC	<dbl>	Orthogonal direction change
MV	<dbl>	Movement variability
ME	<dbl>	Movement error
MO	<dbl>	Movement offset


# Loading data
```{r}

load_data <- function(path) { 
  files <- dir(path, pattern = '\\.sd2', full.names = TRUE)
  tables <- lapply(files, read.csv)
  bind_rows(tables)
}

Participant1 <- load_data("Participant1")

Participant2 <- load_data("Participant2")

Participant3 <- load_data("Participant3")

Participant7 <- load_data("Participant7")

Participant8 <- load_data("Participant8")

Participant9 <- load_data("Participant9")

Participants_raw <- bind_rows(Participant1, Participant2, Participant3, Participant7, Participant8, Participant9)

Participants_raw
```

# Tidying/wrangling 
```{r}


Participants_raw$Participant[Participants_raw$Participant == "P07"] <- "P7"
Participants_raw$Participant[Participants_raw$Participant == "P08"] <- "P8"
Participants_raw$Participant[Participants_raw$Participant == "P09"] <- "P9"
Participants <- Participants_raw
Participants
```

```{r Ensure that independent variables are factors}

Participants$Condition <- as.factor(Participants$Condition)

```



# Plotting beeswarm of the data points and the mean throughput (by participant):
```{r}
pd <- position_dodge(0.3) # ensure no overlaps

plot_grid(
  { 
  Participants %>% 
    ggplot(aes(x = Condition, y = TP.bps., color = Participant)) +
    geom_beeswarm() + 
    ylab("Mean Throughput")+
    expand_limits(y = 0) +
    theme(legend.position = "none", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },{
  Participants %>% 
    ggplot(aes(x = Condition, y = TP.bps., color =  Participant, group = Participant)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 1, position = pd) + 
    ylab("Mean Throughput and 95% CI") +
    expand_limits(y = 0) +
    theme(legend.position = "top", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },
  nrow = 1)
```

# Plotting the mean throughput (by block):
```{r}
pd <- position_dodge(0.3) # ensure no overlaps

plot_grid(
  { 
  Participants %>% 
    ggplot(aes(x = Condition, y = TP.bps., color = Block)) +
    geom_beeswarm() + 
    ylab("Mean Throughput")+
    expand_limits(y = 0) +
    expand_limits(y = 4.25) +
    theme(legend.position = "none", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },{
  Participants %>% 
    ggplot(aes(x = Condition, y = TP.bps., color =  Block, group = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 1, position = pd) + 
    ylab("Mean Throughput and 95% CI") +
    expand_limits(y = 0) +
    expand_limits(y = 4.25) +
    theme(legend.position = "right", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },
  nrow = 1)
```
In general the second block appears to show slightly higher throughput in completing the tasks. In the case of the Laptop condition it appears almost significant, but for the mobile touchpad there is little difference from one block to the next. In the case of Sway the shift is greater, suggesting possible learning effect. 

# Plotting beeswarm of the data points and the mean time (by participant):
```{r}
pd <- position_dodge(0.3) # ensure no overlaps

plot_grid(
  { 
  Participants %>% 
    ggplot(aes(x = Condition, y = MT.ms., color = Participant)) +
    geom_beeswarm() + 
    ylab("Mean Time")+
    expand_limits(y = 0) +
    theme(legend.position = "none", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },{
  Participants %>% 
    ggplot(aes(x = Condition, y = MT.ms., color =  Participant, group = Participant)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 1, position = pd) + 
    ylab("Mean Time and 95% CI") +
    expand_limits(y = 0) +
    theme(legend.position = "top", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },
  nrow = 1)
```
We observe in the above plots the extreme effect of a single outlier sequence on Participant 8 (P8)'s results. 

# Plotting the mean time (by block):
```{r}
pd <- position_dodge(0.3) # ensure no overlaps

plot_grid(
  { 
  Participants %>% 
    ggplot(aes(x = Condition, y = MT.ms., color = Block)) +
    geom_beeswarm() + 
    ylab("Mean Time")+
    expand_limits(y = 0) +
    theme(legend.position = "none", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },{
  Participants %>% 
    ggplot(aes(x = Condition, y = MT.ms., color =  Block, group = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 1, position = pd) + 
    ylab("Mean Time and 95% CI") +
    expand_limits(y = 0) +
    expand_limits(y = 4) +
    theme(legend.position = "top", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },
  nrow = 1)
```
In general the second block is nearly on or contained within the confidence interval of the first block. Hoewver, a slight increase in speed (reduction in mean time) is viewed in the 2-handed touch and the sway conditions. This could be learning effect. 

# Plotting the mean error (by blocks, by participant):
```{r}

plot_grid(
  # By blocks
  {
    Participants %>% 
    ggplot(aes(x = Condition, y = ER..., color =  Block, group = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 1, position = pd) + 
    xlab("") +
    ylab("Mean error percentage") +
    expand_limits(y = 0) +
    expand_limits(y = 20) +
    theme(legend.position = "top", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },
  # By participants
  {
  Participants %>% 
    ggplot(aes(x = Condition, y = ER..., color =  Participant, group = Participant)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    xlab("") +
    ylab("") +
    expand_limits(y = 0) +
    theme(legend.position = "top", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },
  nrow = 1)
```
Mean error does not appear to be significantly affected by block. However, participant 9 in particular had more difficulty on average than other participants.

# Plotting jitter and mean error percentage:
```{r}

plot_grid(
  # By participants
  { 
  Participants %>% 
    ggplot(aes(x = Condition, y = ER..., color = Participant)) +
    geom_jitter() + 
    ylab("Mean error %")+
    expand_limits(y = 0) +
    theme(legend.position = "top", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },{
  Participants %>% 
    ggplot(aes(x = Condition, y = ER...)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 1, position = pd) + 
    ylab("Mean error % and 95% CI") +
    expand_limits(y = 0) +
    theme(axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  },
  nrow = 1)
```
As is most clear in the jitter, the majority of sequences involved 0 errors. With the 95% CI error bars it can be seen the one-handed touch (C02) and the sway (C04) are both generating more errors than the laptop trackpad baseline condition. Two-handed touch appears to be slightly higher in errors than the laptop condition, but not necessarily significantly. 

# Plotting various additional accuracy measures:
Now we look at the four different accuracy measures FittsTwo offers: Error percentage, movement direction change, orthogonal direction change, and movement error. 
```{r}


  # Standard 

p1 <-  Participants %>% 
    ggplot(aes(x = Condition, y = ER..., color = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    xlab("") +
    ylab("ER %") +
    expand_limits(y = 0) +
    theme(legend.position = "none", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))

  # MDC

p2<-  Participants %>% 
    ggplot(aes(x = Condition, y = MDC, color = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    xlab("") +
    ylab("MDC") +
    expand_limits(y = 0) +
    theme(legend.position = "none", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))

  # ODC

p3<-  Participants %>% 
    ggplot(aes(x = Condition, y = ODC, color = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    xlab("") +
    ylab("ODC") +
    expand_limits(y = 0) +
    theme(legend.position = "none", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))

  # ME

p4<-  Participants %>% 
    ggplot(aes(x = Condition, y = MV, color = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    xlab("") +
    ylab("ME") +
    expand_limits(y = 0) +
    theme(legend.position = "none", axis.text.x = element_text(angle=30), axis.title.x=element_blank()) + 
    scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))

legend <- get_legend(p1 + theme(legend.position="bottom"))
 grid <- plot_grid(p1, p2, p3, p4, nrow = 2)
 plot_grid(grid, legend, ncol = 1, rel_heights = c(1, .1))
```
With the exception of ME (mean error) which shows an inflated value for 1-handed touch, the error/accuracy measures have very similar relative distribution across conditions. For blocks all four accuracy measures show fairly tight groupings relative to condition. 


# PART I Analysis: Linear model_1 (with throughput as independent variable)
```{r}
m1 <- lm( TP.bps. ~ Condition, data = Participants)
summary(m1)
anova(m1)
```

```{r}
m1$coefficients
```

# Plotting the results of the linear model_1

```{r} 

coeffs_df <- tibble(
  Coeff_id = paste0("b", 0:(length(m1$coefficients)-1)),
  Coeff = m1$coefficients,
  Plot_x = levels(Participants$Condition),
  Plot_y = c(m1$coefficients[1], m1$coefficients[1], m1$coefficients[1], m1$coefficients[1]),
  Plot_yend = c(m1$coefficients[1], 
            m1$coefficients[1] + m1$coefficients[2], 
            m1$coefficients[1] + m1$coefficients[3],
            m1$coefficients[1] + m1$coefficients[4])
)

Participants %>% 
  ggplot(aes(x = Condition, y = TP.bps.)) +
  geom_beeswarm(alpha = 0.3) +
  geom_point(data = coeffs_df, 
             aes(x = Plot_x, y = Plot_yend), 
             color = "red", shape = 18, size = 3) +
  geom_segment(data = coeffs_df, 
               aes(x = Plot_x, xend = Plot_x, y = Plot_y, yend = Plot_yend), 
               color = "red",
               arrow = arrow(length = unit(0.03, "npc"))) +
  geom_hline(yintercept = m1$coefficients[1], color = "red", linetype = "dashed") +
  ylab("Throughput (bps)") +
  scale_x_discrete(labels = c("Laptop", "1-hand touch", "2-hand touch", "Sway"))
  

```
In the above model the baseline condition (C01: laptop) is being treated as the y-intercept, with 0 slope. This is useful for visually illustrating the decreased throughput for each of the remote mouse conditions, but is not a very useful predictive model (presumably because the conditions are factors, and do not have a defined logical ordering as a numeric metric or Block# would). 


# Hypotheses and planned contrast regarding throughput
* \\(H_1\\): Mobile phone based “remote mouses” have a higher throughput than laptop based touchpads. (C02+C03+C04)-C01=0
* \\(H_3.2\\): Mobile phone based “remote mouses” have a higher throughput when using two-handed pointing method compared to one-handed pointing method. C03-C02=0


```{r}
  m1.contrast <- glht(m1, linfct = mcp(Condition =
  c("(C02+C03+C04)-C01=0",
    "C03-C02=0"
    )))

  m1.contrast %>%
  confint() %>%
  tidy() %>%
  ggplot(aes(x = paste(lhs, "==", rhs))) +
  geom_hline(yintercept = 0, color = "red") +
  geom_point(aes(y = estimate)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0) +
  expand_limits(y = 0) +
  coord_flip() +
  xlab("Hypothesis") +
  ylab("Estimate of the difference") 
  summary(m1.contrast)
```

**Interpretation:** Recall: the nearer to the zero, the more valid is the hypothesis 
H_1) The estimate of the difference between mobile phone based “remote mouses” and laptop based touchpad  is roughly -2.68, what means the hypothesis should not be accepted, since throughput with laptop based touchpad is substantially larger compared to all mobile phone based conditions

H_2) The estimate of the difference between two-handed pointing method and one-handed pointing method on the mobile phone is close to zero (+0.22). The throughput in two-handed method is only minimal larger, which may be in this case not enough to support the hypothesis.



#Part II Analysis: 

## Automatic model comparison: dependence of the throughput from 7 accuracy measures
```{r}
lm_all <- lm(
  TP.bps. ~  
  MT.ms. +         # Speed represented by movement time
  TRE +            # 1. Target re-entry
  TAC +            # 2. Task axis crossing
  MDC +            # 3. Movement direction change
  ODC +            # 4. Orthogonal direction change
  MV +             # 5. Movement variability
  ME +             # 6. Movement error
  MO,              # 7. Movement offset
        
  data = Participants   
  )


step_backward_result <- 
  MASS::stepAIC(lm_all, 
    direction = "backward", 
    trace = TRUE           # this shows the fit trace so we can track AIC (optional)
    )

step_backward_result$anova
anova_result <- anova(step_backward_result)
print(anova_result)
```


**Interpretation:** Throughput is an accepted measure — now endorsed in an ISO standard (MacKenzie et al (2001), p. 14). It has two core components: (a) speed, represented by movement time and (b) accuracy, represented by 7 accuracy measures, mentioned above. Results from model comparison still retain MT(movement time) as one of the core components emphasize 3 from 7 accuracy measures:   `MDC`(Movement direction change), `ODC`(Orthogonal direction change) and `ME`(Movement error), which have the greatest impact on the linear model and result in the lowest AIC score based on our experiment data. Therefore in the further analysis we will concentrate on this 3 accuracy measures to explain the differences between the conditions.


