---
title: "Assignment 3: Analysis of pointing methods"
output:
  pdf_document: default
  html_notebook:
    code_folding: hide
  html_document: default
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE)


####
# BEGIN: Chat's R header (revision: 10.4.17)
####

# The "pacman" package allows p_load() and p_load_gh() which
# automatically install missing packages and load them
if (!require("pacman")) install.packages("pacman", repos='https://stat.ethz.ch/CRAN/'); library(pacman)

p_load(
   car,        # grab bag of useful functions for NHST
   multcomp,   # for glht (planned comparison, post-hoc test)
   tidyverse,  # collection of the tidyverse packages (this automatically load the following):
   #dplyr,     #   - for data wrangling
   #tibble,    #   - a stricter alternative to data.frame
   #readr,     #   - a stricter alternative to read.csv
   #ggplot2,   #   - for plotting
               # other packages in tidyverse that are non-core
   stringr,    #   - for string functions
   tidyr,      #   - for data tidying
   forcats,    #   - utility functions for working with factor levels
               # extensions of tidyverse
   broom,      #   - for cleaing output from models, e.g., lm()
   cowplot,    #   - adds plot_grid() to put multiple ggplot()'s togeter
   GGally,     #   - adds ggpairs() which is a smarter scatterplot matrix
               # data structure
   GetoptLong, #   - string interpolation. See qq() explanation below
   lubridate,  #   - utility for parsing and performing arithematic on dates 
               # 
               # visualization & interactivity
   Hmisc,      #   - for plotting mean and CI in ggplot
   rafalib,    #   - for imagemat function (visualize contrast codings)
   DT,         #   - for showing data table with navigation/search controls
               # testing:
   assertthat  #   - unit-testing functions
)

p_load_gh(
   "eclarke/ggbeeswarm" # beeswarm plot extension for ggplot2
)

# GetoptLong config (string interpolation)
qq.options("code.pattern" = "#\\{CODE\\}") 

# ggplot2 config (plot theme)
myTheme <- theme(
   panel.background = element_blank(), 
   panel.grid.major = element_line(color="lightgrey", size = 0.2)
)

# DT config
options(DT.options = list(pageLength = 10))
options(DT.autoHideNavigation = TRUE)

# Optional: Decimal output readability
# options(digits=2)
##   NOTE: This option might cause output to be printed with rounding. (default value = 7)

####
# END: Chat's R header
####
```

<!-- two columns -->
<style>
.column-left{
  float: left;
  width: 50%;
  text-align: left;
  padding-right: 1%;
}
.column-right{
  float: right;
  width: 50%;
  text-align: left;
  padding-left: 1%;
}
.clear {
  clear: both;
}
</style>

# Data and exploratory plot
The following dataset is simulated to be similar to the comparison of tangible knobs and touch input (Voelker et al., 2015). In this dataset, we simplify the study with the following setup:

* 2 Devices (Touch and Knob) x
* 3 Vision conditions (EyesFree, Peripheral, and EyesOn)
* Between-subjects design
* The dependent variable is the completion time in seconds.
* No repetitions
```{r}

load_data <- function(path) { 
  files <- dir(path, pattern = '\\.sd2', full.names = TRUE)
  tables <- lapply(files, read.csv)
  bind_rows(tables)
}

Participant1 <- load_data("Participant1")

Participant2 <- load_data("Participant2")

Participant3 <- load_data("Participant3")

Participant7 <- load_data("Participant7")
Participant7$Participant = "P7"

Participant8 <- load_data("Participant8")
Participant8$Participant = "P8"

Participant9 <- load_data("Participant9")
Participant9$Participant = "P9"

Participants <- bind_rows(Participant1, Participant2, Participant3, Participant7, Participant8, Participant9)

Participants

```

```{r}
# ensure that independent variables are factors (essential for glht())
Participants$Condition <- as.factor(Participants$Condition)
```

Plotting beeswarm of the data points and the mean time:

```{r}
pd <- position_dodge(0.3) # ensure no overlaps

plot_grid({ 
  Participants %>% 
    ggplot(aes(x = Condition, y = MT.ms., color = Block)) +
    geom_beeswarm() + 
    ylab("Mean time (ms)") +
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  Participants %>% 
    ggplot(aes(x = Condition, y = MT.ms., color = Block, group = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    labs(title = "By Block", size=15) +
    expand_limits(y = 0)
  },{
   Participants %>% 
    ggplot(aes(x = Condition, y = MT.ms., color = Participant)) +
    geom_beeswarm() + 
    ylab("") +
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  Participants %>% 
    ggplot(aes(x = Condition, y = MT.ms., color = Participant, group = Participant)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("") +
    labs(title = "By Participant", size=15) +
    expand_limits(y = 0)
  }, 
  nrow = 2)
```

**Interpretation:** The plot on the right shows the mean and 95% confidence interval of each group. There seems to be an overall effect of `Device`: knob seems to be slower than touch. As for `Vision`, there seems to be an effect: eyes-on condition is fastest, follow by the peripheral condition, and the eyes-free condition is slowest. The slope of the three lines seems to differ, suggesting an interaction between `Device` and `Vision`.


Plotting mean time for each condition, grouped by participant:

```{r}
pd <- position_dodge(0.2) # ensure no overlaps

  Participants %>% 
    ggplot(aes(x = Condition, y = MT.ms., color = Participant, group = Participant)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)

```
As one can see in the above plot, extreme outliers have a significant impact on both the participant and overall (group) results for participant 8 (C04), and to a lesser degree participant 9 (C03). 

C01 (laptop) is fastest for all participants, and in general C03 (two-handed) was faster than C02 (one-handed). However, results for C04 (sway) varied significantly between participants. 

Plotting effects of amplitude and width on mean time, grouped by participant
```{r}
pd <- position_dodge(0.3) # ensure no overlaps

plot_grid({ 
  Participants %>% 
    ggplot(aes(x = A, y = MT.ms., color = Participant)) +
    geom_beeswarm() + 
    ylab("Mean time (ms)") +
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  Participants %>% 
    ggplot(aes(x = A, y = MT.ms., color = Participant, group = Participant)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)
  },{
   Participants %>% 
    ggplot(aes(x = W, y = MT.ms., color = Participant)) +
    geom_beeswarm() + 
    ylab("Mean time (ms)") +
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  Participants %>% 
    ggplot(aes(x = W, y = MT.ms., color = Participant, group = Participant)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)
  }, 
  nrow = 2)

```
Plotting effects of amplitude and width on mean time, grouped by block
```{r}
pd <- position_dodge(0.3) # ensure no overlaps

plot_grid({ 
  Participants %>% 
    ggplot(aes(x = A, y = MT.ms., color = Block)) +
    geom_beeswarm() + 
    ylab("Mean time (ms)") +
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  Participants %>% 
    ggplot(aes(x = A, y = MT.ms., color = Block, group = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)
  },{
   Participants %>% 
    ggplot(aes(x = W, y = MT.ms., color = Block)) +
    geom_beeswarm() + 
    ylab("Mean time (ms)") +
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  Participants %>% 
    ggplot(aes(x = W, y = MT.ms., color = Block, group = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)
  }, 
  nrow = 2)

```


# Removing extreme outliers (2) and retesting

```{r}

Participants <- filter(Participants, MT.ms. <= 6000)

plot_grid({ 
  Participants %>% 
    ggplot(aes(x = A, y = MT.ms., color = Participant)) +
    geom_beeswarm() + 
    ylab("Mean time (ms)") +
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  Participants %>% 
    ggplot(aes(x = A, y = MT.ms., color = Participant, group = Participant)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)
  },{
   Participants %>% 
    ggplot(aes(x = W, y = MT.ms., color = Participant)) +
    geom_beeswarm() + 
    ylab("Mean time (ms)") +
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  Participants %>% 
    ggplot(aes(x = W, y = MT.ms., color = Participant, group = Participant)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)
  }, 
  nrow = 2)

```
```{r}
pd <- position_dodge(0.3) # ensure no overlaps

plot_grid({ 
  Participants %>% 
    ggplot(aes(x = A, y = MT.ms., color = Block)) +
    geom_beeswarm() + 
    ylab("Mean time (ms)") +
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  Participants %>% 
    ggplot(aes(x = A, y = MT.ms., color = Block, group = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)
  },{
   Participants %>% 
    ggplot(aes(x = W, y = MT.ms., color = Block)) +
    geom_beeswarm() + 
    ylab("Mean time (ms)") +
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  Participants %>% 
    ggplot(aes(x = W, y = MT.ms., color = Block, group = Block)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)
  }, 
  nrow = 2)

```

# Data and exploratory plot
The interaction effect is coded by multiplying regressors. Notice the comparison between the two model matrices below. On the left matrix, no interaction effect is coded. Thus, we have four regressors: one intercept, one regressor for `Device`, two regressors for  `Vision`. On the right matrix, two regressors are added to encode the interactions between `Device` and `Vision`. They are the result of multiplication between the second regressor and the third (and fourth respectively).

```{r}
par(mfcol = c(1,2))
rafalib::imagemat(model.matrix(~ Device + Vision, data = data1), main ="Without interaction")
coded2 <- model.matrix(~ Device * Vision, data = data1) 
rafalib::imagemat(coded2, main ="With interaction")
par(mfcol = c(1,1))
```
# Model fitting
```{r}
m1 <- lm(MT.ms. ~ Condition, data = Participants)
summary(m1)
anova(m1)
```
```{r}
m2 <- lm(MT.ms. ~ W, data = Participants)
summary(m2)
anova(m2)
```


# Summary
For categorical independent variables with interaction effects, we analyze them using the following steps:

1. Fit model with interaction to assess the effect of the interaction term.

Then, follow up with planned contrasts (preferred) or multiple comparisons of:

2. Compare means of the main efffects (the effect of each independent variable). Here, you have to remove the interaction effect term before running `ghlt()`

3. Compare means of the interaction effects. Here, you have to  remove the main effect terms as well as the intercept before running `ghlt()`


It's possible to do only 2. or 3., depending on your hypotheses.


# Your turn

The simulated dataset below comes from a study comparing touch reaction time. The participants tap on a target that appears on the screen. The study is between-subjecs design with two independent variables:

* `Device`: two sizes of mobile devices (`Phone` vs. `Tablet`)
* `Hands`: whether the participants use only the index finger of their dominant hand (`OneIndex`) or using two thumbs from both hands (`TwoThumbs`)

You are interested in the effect of `Devices` and `Hands` as well as their interactions. Analyze this dataset by completing the code below

## Data loading & exploratory plot
```{r}
data_typing <- read_csv("../data/TouchscreenTyping.csv", col_types = "cccd")
data_typing %>%
  filter(., Hands == "TwoThumbs" & Device == "Tablet") %>%
  mutate(Time = Time - 0.5) ->
  data_typing[data_typing$Hands == "TwoThumbs" & data_typing$Device == "Tablet", ]

data_typing %>%
  filter(., Hands == "TwoThumbs" & Device == "Phone") %>%
  mutate(Time = Time + 0.5) ->
  data_typing[data_typing$Hands == "TwoThumbs" & data_typing$Device == "Phone", ]

pd <- position_dodge(0.3) # ensure no overlaps

plot_grid({ 
  data_typing %>% 
    ggplot(aes(x = Device, y = Time, color = Hands)) +
    geom_beeswarm() + 
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  data_typing %>% 
    ggplot(aes(x = Device, y = Time, color = Hands, group = Hands)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)
  },
  nrow = 1)
```

**Interpretation:** TODO: write your preliminary interpretation based on the plot above.

## Assess model fit with interaction term

```{r}
# # ensure data type for modelling
data_typing$Device <- as.factor(data_typing$Device)
data_typing$Hands <- as.factor(data_typing$Hands)

# # linear model and ANOVA
m2 <- lm(Time ~ Hands * Device, data = data_typing)
anova(m2)
```

**Interpretation:** TODO: Write your interpretation of the model fit. (Hint: Which terms seems to have a significant effect according to ANOVA result?)

## Main effect comparison


```{r}
m2.main <-  update(m2, .~. - Hands:Device)   # update the model to exclude the interaction
m2.main.pairwise <- glht(m2.main,
  linfct = mcp(
    Device = "Tukey",
    Hands = "Tukey"
    ))
print(m2.main.pairwise)
m2.main.pairwise %>% plotMc()
```

**Interpretation:** TODO: Write your interpretation. (Hint: Which independent variables seems to have stronger effect?)

The difference from two thumbs to one index is greater than the difference from tablet to phone. 

## Interaction comparison
Do multiple comparison to address the hypotheses below:

* \\(H_{0, 1}\\): On the phone, there's no difference between the two-thumbs and one-index usage.
* \\(H_{0, 2}\\): On the tablet, there's no difference between the two-thumbs and one-index usage.

```{r}
m2.interaction <-  update(m2, .~. - Device - Hands -1)   # update the model to exclude main effect and intercept
summary(m2.interaction)
```

In the contrast matrix below, the columns are the levels of the first variable of the interaction

```{r}
# # template of the contrast matrix
contr_matrix <- matrix(byrow = TRUE, nrow = 2,
  data = c(
    0, 0, 0, 0,
    0, 0, 0, 0),
  dimnames = list(c(
      # row names: hypotheses (only the left-hand-side)
      "Phone: TwoThumbs - OneIndex",
      "Tablet: TwoThumbs - OneIndex"
    ),
    c(
      # column names: the levels of the first variable of the interaction
      rep(m2.interaction$xlevels[[1]], 2)
    )))

print(contr_matrix)

# # TODO: replace some of the following zeros with an appropriate number. 
contr_matrix[1,] <- c(-1, 1, 0, 0)
contr_matrix[2,] <- c(0, 0, -1, 1)
print(contr_matrix) # updated contrast matrix
```

Now, you can run the interaction contrasts.
```{r}
m2.interaction.mc <- glht(m2.interaction, linfct = contr_matrix)
m2.interaction.mc %>%plotMc()
```

**Interpretation:** TODO: Write your interpretation. (Hint: Does the effect of the hands manifest in the same way in both devices?)

The effect of the hands manifests in both devices, but to a much lesser extent on the phone. For tablet it makes a much more significant difference. 


## Your turn (2)

You replicated the experiment above with an addition of a smartwatch condition. Analyze the data from this experiment.

```{r}
data_typing2 <- read_csv("../data/TouchscreenTyping2.csv", col_types = "cccd")

plot_grid({ 
  data_typing2 %>% 
    ggplot(aes(x = Device, y = Time, color = Hands)) +
    geom_beeswarm() + 
    expand_limits(y = 0) +
    theme(legend.position = "none")
  },{
  data_typing2 %>% 
    ggplot(aes(x = Device, y = Time, color = Hands, group = Hands)) +
    stat_summary(fun.y = mean, geom = "point", shape = 18, size = 3, position = pd) + 
    stat_summary(fun.y = mean, geom = "line", position = pd) + 
    stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0, position = pd) + 
    ylab("Mean(Time) and 95% CI") +
    expand_limits(y = 0)
  },
  nrow = 1)
```

```{r}
# # ensure data type for modelling
data_typing2$Device <- as.factor(data_typing2$Device)
data_typing2$Hands <- as.factor(data_typing2$Hands)

# # linear model and ANOVA
m3 <- lm(Time ~ Hands * Device, data = data_typing2)
anova(m3)

m3.main <-  update(m3, .~. - Hands:Device)   # update the model to exclude the interaction
m3.main.pairwise <- glht(m3.main,
  linfct = mcp(
    Device = "Tukey",
    Hands = "Tukey"
    ))
print(m3.main.pairwise)
m3.main.pairwise %>% plotMc()

m3.interaction <-  update(m3, .~. - Device - Hands -1)   # update the model to exclude main effect and intercept
summary(m3.interaction)

# # template of the contrast matrix
contr_matrix2 <- matrix(byrow = TRUE, nrow = 3,
  data = c(
    0, 0, 0, 0, 0, 0,
    0, 0, 0, 0, 0, 0,
    0, 0, 0, 0, 0, 0),
  dimnames = list(c(
      # row names: hypotheses (only the left-hand-side)
      "Phone: TwoThumbs - OneIndex",
      "Tablet: TwoThumbs - OneIndex",
      "Watch: TwoThumbs - OneIndex"
    ),
    c(
      # column names: the levels of the first variable of the interaction
      rep(m3.interaction$xlevels[[1]], 3)
    )))

print(contr_matrix2)

# # TODO: replace some of the following zeros with an appropriate number. 
contr_matrix2[1,] <- c(-1, 1, 0, 0, 0, 0)
contr_matrix2[2,] <- c(0, 0, -1, 1, 0, 0)
contr_matrix2[3,] <- c(0, 0, 0, 0, -1, 1)
print(contr_matrix2) # updated contrast matrix

m3.interaction.mc <- glht(m3.interaction, linfct = contr_matrix2)
m3.interaction.mc %>%plotMc()

```

# References
* Simon Voelker, Kjell Ivar Øvergård, Chat Wacharamanotham, and Jan Borchers. 2015. Knobology Revisited: A Comparison of User Performance between Tangible and Virtual Rotary Knobs. In Proceedings of the 2015 International Conference on Interactive Tabletops & Surfaces (ITS '15). ACM, New York, NY, USA, 35-38. DOI: https://doi.org/10.1145/2817721.2817725